---
title: "Lab 02 - Plastic waste"
author: "Insert your name here"
date: "Insert date here"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises


### Exercise 1

```{r plastic-waste-continent}
ggplot(data = plastic_waste,mapping = aes(x = plastic_waste_per_cap, color = continent, fill = continent))+
geom_density(alpha = 0.7)
```

### Exercise 2

```{r plastic-waste-density}
#2.1
ggplot(data = plastic_waste,mapping = aes(x = plastic_waste_per_cap, color = continent, fill = continent))+
geom_density(alpha = 0.1)
#2.2
ggplot(data = plastic_waste,mapping = aes(x = plastic_waste_per_cap, y = continent))+
geom_boxplot()
```

### Exercise 3
Violin plot reveals the frequency of plastic waste per cap. We can better see the concentration of the same plastic waste per cap better compared to box plot. 

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
ggplot(data = plastic_waste,mapping = aes(x = continent , y = plastic_waste_per_cap))+
geom_violin()
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
data <- plastic_waste %>%
  filter(plastic_waste_per_cap <3.5)

ggplot(data = data, aes(x = mismanaged_plastic_waste_per_cap , y = plastic_waste_per_cap))+
geom_point()
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
ggplot(data = plastic_waste,mapping = aes(x = mismanaged_plastic_waste_per_cap , y = plastic_waste_per_cap, color= continent))+
geom_point()
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
ggplot(data = plastic_waste,mapping = aes(x = total_pop , y = plastic_waste_per_cap, color= continent))+
geom_point()

cor(plastic_waste$plastic_waste_per_cap, plastic_waste$total_pop, use = "pairwise.complete.obs")
```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste,mapping = aes(x = coastal_pop , y = plastic_waste_per_cap, color= continent))+
geom_point()

cor(plastic_waste$plastic_waste_per_cap, plastic_waste$coastal_pop, use = "pairwise.complete.obs")
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
data <- plastic_waste %>%
  filter(plastic_waste_per_cap < 3)
data$y <- data$coastal_pop/data$total_pop

ggplot(data = data, aes(x = y, y = plastic_waste_per_cap))+
  geom_point(mapping = aes(color=continent))+
  geom_smooth()+
  labs(title= "plastic_waste v.s coastal population proportion",
       subtitle = "by continent",
       x = "plastic waste per capita",
       y = "coastal population proportion (coastal/total population)")+
  scale_color_viridis_d()
  
rm(data)
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



